During specification development, you will likely encounter situations where the 
specification is not realizable. Salty can help a little bit with this, 
by identifying a few situations that will cause slugs to reject your specification, 
before realizability is checked.

\section{Unsatisfiable Safety Constraints}

As you can specify safety constraints on the environment and system in 
many different places, it's easy to end up with an unsatisfiable specification. 
As an example, consider the following specification:

\begin{lstlisting}
controller Example where

input  x1: Bool
input  x2: Bool
output y1: Bool
output y2: Bool

env_trans   x1
env_trans ! x1

env_trans x2' == !x2

sys_trans   y1
sys_trans ! y1

sys_trans y2' == !y2
\end{lstlisting}
The sanity checker in salty will reject this specification for two reasons:
\begin{enumerate}
 \item The environment safety properties are unsatisfiable. This by itself isn't a problem 
for slugs, as the semantics it uses allow it to produce a controller, as long as the 
system satisfies its safety properties for at least one step. However, the sanity 
checker in salty rejects this situation, as it produces a controller that would 
satisfy system safety for exactly one step, and then behave unpredictably forever.
 \item The system safety properties are unsatisfiable. This is a problem, as slugs will 
refuse to synthesize a controller for a specification that will never be able to 
satisfy its safety properties.
\end{enumerate}

Salty will output two errors in response to these problems:
\begin{lstlisting}
-- [error] --------------------------------------------------
  Environment safety constraints are never satisfiable:
    * environment safety constraint at example.salt:8,13-8,15
    * environment safety constraint at example.salt:9,11-9,15

-- [error] --------------------------------------------------
  System safety constraints are never satisfiable
    * system safety constraint at example.salt:14,11-14,15
    * system safety constraint at example.salt:13,13-13,15
\end{lstlisting}
The errors attempt to report a minimal set of conflicting safety properties, 
to aid specification debugging; the errors above don't mention the safety properties 
on lines 11 and 15 because they are not contributing to the unsatisfiability of 
the environment or system.

\section{Unsatisfiable Initial Constraints}

Once salty has managed to determine that the safety properties of the system are 
satisfiable, it will continue by checking the initial state of the controller. 
If any of the initial value constraints for state variables cause the environment 
or system safety properties to become unsatisfiable, an error is raised.

Although slugs would be able to synthesize a controller whose variable initialization 
violated environmental safety properties, salty considers that to be an error. 
The reasoning for this is the same as when checking environmental safety properties 
for satisfiability: if the controller produced would only satisfy its system safety 
properties for a single time step, that is not a useful controller.

\section{Unsafe Liveness Properties}

Liveness properties in the system and environment can eventually cause safety 
properties to be violated. Consider this example specification:
\begin{lstlisting}
controller Example where

input a : Bool = False

env_trans ! a

env_liveness a
\end{lstlisting}
This specification will generate a warning, because it will eventually violate its 
environmental safety properties. The reason that this isn't rejected is that the 
semantics of slugs requires that the system be able to satisfy its safety properties 
during the state transition when the environment violates its own. All states after 
that are allowed to violate system safety.

If we augment the same example with a system safety property that will be violated 
by a liveness property, this will also produce a warning, as salty is not able to 
tell if the environment or the system would violate safety first. As a result, this 
specification will still be sent to slugs, and may be unrealizable.

However, if there is a liveness property that would cause a system safety violation, 
and no environment safety violation can be established, the sanity checker generates 
an error and salty rejects the specification.
